---
title: "Gov 50 Final Project"
author: Michelle Amponsah 
description: Effect of Source of News Information on Thoughts on Democracy  
output:
  distill::distill_article:
    self_contained: false
---

In this project I seek to study the relationship between the source of news 
information in countries on the African continent and its effect on participants’
support for democracy. In other words, the research question I will answer is 
whether or not where one consumes their news (from traditional newspapers, 
internet, or social media platforms such as Facebook or Twitter) affects 
support for democracy. I hypothesize that those who more often receive their news 
from social media sites like Facebook and Twitter have weaker support for democracy
because of the accessibility of extreme viewpoints on social media sites, “echo
chambers” that reinforce these viewpoints, and misinformation. I hypothesize 
that those who receive their news from the internet tend to support democracy 
more than those who receive their news from social media but less than those who 
receive the majority of their news from newspapers as digital news information 
(non-social media) encompasses a wide spectrum of viewpoints. 

The data is taken from the responses from a survey administered by Afrobarometer,
a non-partisan and pan-African research network that seeks to measure attitudes 
surrounding socio-politics and economics on the African continent. The 
explanatory variable is question 30, which asks respondents to identify which of 
three statements is closest to their own opinion. The statements are: 

`Statement 1: Democracy is preferable to any other kind of government.`
`Statement 2: In some circumstances, a non-democratic government can be preferable.`
`Statement 3: For someone like me, it doesn’t matter what kind of government we have.`

The variable is coded such that 1 = Statement 3, 2 = Statement 2, and 3 = Statement
1. Again, 9 = don't know, 98 = refused to answer, and -1 = missing. To identify 
the relationship between a specific news sources and thoughts on democracy, I'd 
have to conduct separate visualizations such that one news source is on the x
axis and thoughts on democracy is on the y-axis. A negative correlation between
the two variables would indicate that the news source is correlated to weaker opinions
on democracy while a positive correlation would indicate the news source is correlated
to higher esteem of democracy. 


```{r}
library(tidyverse)
library(dplyr)
library(readr)
load("/Users/michelleamponsah/Downloads/afrobarometer2016.RData")
afro <- afro |> 
select(RESPNO, Q12A:Q12E, Q30)

afro
```

```{r}
library(gov50data)
afro <- afro |>
  rename(radio_news = Q12A,
         tv_news = Q12B,
         newspapers = Q12C,
         internet = Q12D,
         social_med = Q12E,
         democracy_support = Q30)
afro
```


```{r}
afro <- afro |> 
  mutate(support_dem = case_when(democracy_support == "1" ~ "Statement 3",
                                 democracy_support == "2" ~ "Statement 2",
                                 democracy_support == "3" ~ "Statement 1"))

afro
```

##visualization 1: 


```{r}
afro <- afro |> 
  drop_na(tv_news, support_dem)


```

```{r}

tv_news_hist <- afro |> 
  group_by(support_dem) |>
  ggplot(data = afro, mapping = aes(x = tv_news,
                                    y = support_dem)) + geom_col(
                                      mapping = aes(color = support_dem)) +
  labs(x = "Frequency Respondent Receives News from TV",
       y = "Support for Democracy",
       title = "TV News and Democracy Support") + scale_x_log10() 

tv_news_hist

```