---
title: "Gov 50 Final Project"
author: Michelle Amponsah 
description: Effect of Source of News Information on Thoughts on Democracy  
output:
  distill::distill_article:
    self_contained: false
---


# Introduction 


I am a newsie. I love the news — one of the first clubs I joined my freshman 
fall was The Crimson, and ever since I've thrown myself into the news cycle. I'm
interested in this question of where people get their news and how it affects
their perception towards things like goverment, democracy, and the world they live
in. In this project I seek to study the relationship between the frequency that
one consumes their news from social media and its effect on their support for 
democratic government. 

My research question is this: does primarily receiving one's news from social
media cause lessened support for democracy? 

I hypothesize that those who more often receive their news 
from social media sites like Facebook and Twitter have weaker support for 
democracy because of the accessibility of extreme viewpoints on social media 
sites, “echo chambers” that reinforce these viewpoints, and misinformation.
The null hypothesis, then, would be that frequent consumption of news from
social media has no effect on support for democracy. 


# The Dataset 

I'll be using data from responses to the Afrobarometer 2018 survey. 
Afrobarometer is a non-partisan, pan-African research organization that
administers the Afrobarometer, "temperature check" survey to measure
attitudes surrounding democracy and governance on the African continent. 
I'll be using responses that pertain to the news consumption part of this survey.
This study is cross-sectional because it takes a "snapshot" of the population
at a particular point in time and the data is observational (I had no influence
or assigning role on the participants' responses). 

For my explanatory variable, I focus on Q12E. This represents the question that 
asks how frequently the respondent gets their news from social media sites like
Facebook or Twitter. 

This question is coded such that 0 = Never, 1 = less than once a month,
2 = a few times a month, 3 = a few times a week, and 4 = every day. 
For the purposes of this study, I am concerned with those who receive their news
from that specific news source from never to everyday (0-4). 

The outcome variable I focus on is Question 30 (`Q30`), which asks respondents
to identify which of the following three statements is closest to their opinion. 

The statements are: 

Statement 1: Democracy is preferable to any other kind of government.
Statement 2: In some circumstances, a non-democratic government can be preferable.
Statement 3: For someone like me, it doesn’t matter what kind of government we have.

The variable is coded such that 1 = Statement 3, 2 = Statement 2, and 3 = Statement
1. Again, 9 = don't know, 98 = refused to answer, and -1 = missing. 

The dataset as it came was hard to read. Therefore, I renamed the question 
variables so that they reflected the type of news they asked respondents about: 


# Loading Afro 

```{r}
library(tidyverse)
library(dplyr)
library(readr)
load("/Users/michelleamponsah/Downloads/afrobarometer2016.RData")
afro <- afro |> 
select(RESPNO, Q12A:Q12E, Q30)

```
## Renaming the Variable 
```{r}
library(gov50data)
afro <- afro |>
  rename(social_med = Q12E) |> 
  filter(social_med == c(0, 1, 2, 3, 4))
```

I also wasn't a fan of the way the `Q30` variable was coded (super confusing!)
so I re-coded it such that it made more sense to me and was a string instead 
of numbers:


# Re-coding democracy_support 

```{r}

library(dplyr)

afro |> 
  count(Q30)


afro <- afro |>
  drop_na(Q30) |> 
    mutate(support_dem = case_when(Q30 == "1" ~ "Statement 3",
                                 Q30 == "2" ~ "Statement 2",
                                 Q30 == "3" ~ "Statement 1"))

```

To identify the relationship between specific social media as a news source
and thoughts on democracy, I will conduct a visualization and regression such that I
can investigate the relationship between frequency of news consumption from social
media and support for democracy. In terms of visualization, a negative 
correlation between the two variables would indicate that the social media news
consumption is correlated to weaker opinions on democracy while a positive 
correlation would indicate the social media news consumption is correlated to 
higher support for democracy. Let's see the distribution of social media news
consumption, first of all: 

# Visualizations and Summary of Social Media News Consumption  

```{r}
afro |>
  ggplot(mapping = aes(x = social_med, y = support_dem)) +
  geom_col(mapping = aes(fill = support_dem)) + labs(x = "Social Media News",
                    y = "Support for Democracy",
                    title = "Distribution of Support for Democracy by
                    Social Media Variable") + scale_x_log10()
```
# Social Media News Summary 
```{r}
soc_news_summary <- afro |>
  group_by(support_dem) |>
  summarize(
    avg_social = mean(social_med),
    sd_social = sd(social_med)
  )

knitr::kable(soc_news_summary,
             col.names = c("Democracy",
                           "Avg Social Media",
                           "SD Social Media"),
             digits = 2)

```

My dependent variable is support for democracy. As the graph shows, as consumption
of news from social media increases, so does likelihood that respondent agrees
with statement 2 and statement 1. 

In the summary (formatted knitted table) we see the average value and standard
deviation of social media news consumption for each statement (the support for
democracy variable).The average value of social media news consumption for
statement 1 was 0.9, for statement 2 was 1.12, and for statement 3 was 0.75 
Interpreted, this means that the highest average consumption of social media
as news was observed for those who chose statement 2, that "In some circumstances,
a non-democratic government can be preferable." In this table, I'm not seeing
a clear relationship emerge between support for democracy and average social
media news consumption. 

# Single Variable Regression 

Since I'm primarily interested in social media news consumption and support for
democracy, my regression will include only that variable. I used `Q30` instead of
`support_dem`because I'll need the numeric values to conduct this regression: 

```{r}

sing <- lm(Q30 ~ social_med, data = afro)

varnames1 <- c("(Intercept)" = "Intercept",
               "social_med" = "Social Media Consumption")

modelsummary::modelsummary(sing, statistic = "p.value", 
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_map = varnames1)
```
As can be seen in the table above the intercept is 2.596, meaning that when
there is no social media news consumption, the support for democracy value is 
at 2.596. The coefficient for social media news consumption, or the slope of the 
line, is 0.008, meaning that for the variable social media consumption, there is 
a .008 increase in support for democracy as social media news consumption
increases. The p-value is 0.099, or 0.1, which is not statistically significant
for an alpha value of .05. Statistical significance would occur if the p-value
was less than .05. Because of the cross-sectional nature of the survey, 
I do believe the coefficient has a causal effect and we can make a causal
interpretation. 


# Summary Plot 

```{r}

afro |>
  ggplot(aes(x = social_med, fill = support_dem)) +
  geom_bar(position = "dodge") +
  labs(x = "Frequency of Social Media Consumption", y = "Count",
       title = "Support for Democracy by Frequency of Social Media Consumption")

```

Based on this barplot, we can observe that support for statement 1
("Democracy is preferable to any other kind of government") is highest when 
the respondents marked that they "never" consume news from social media (0). 
However, as frequency of news consumption from social media consumption increases,
support for statement 1 drops dramatically then increases incrementally. The same
is true for statements 1, 2, and 3, but is most magnified for statement 1.
It appears from the graph that despite frequency of news consumption, respondents
are more likely to support statement 1 than the other two statements. Support
for statements 1, 2, and 3 is lowest when the respondents consumed news from
social media less than once a month. 

# Conclusion 

Based on this study, I fail to reject the null hypothesis of no relationship
between social media news consumption and support for democracy. There is not
a statistically significant relationship between frequency of social media as 
news consumption and support for democratic governance. My hypothesis, then, is
disproved. Further, the coefficient for social media news consumption was very
close to 0, indicating a very weak (but slightly positive) relationship between 
the variables of social media news consumption and support for democracy. Still,
there are limitations to this report. To specify, there may be confounding
variables that complicate the conclusions I draw between social media news
consumption and support for democracy. For example, data like the age, racial
identity, and wealth of the respondent may also affect what they choose for the
support for democracy variable. Further, the answers to other questions about 
news consumption may affect or be conditional upon getting the majority of one's
news from social media; for example, it might be the case that people more likely
to frequently consume news from social media are also more likely to consume
news from the internet, and that is an interesting relationship to explore as 
well. Therefore, there is a case to be made that only studying social media
news consumption does not present the most complete picture of source of news
consumption and support for democracy. Finally, missingness is a significant 
part of the Afrobarometer survey. For the Q30 question about democracy support
alone, there are 1112 missing values (before I dropped them) which presents an 
issue for statistical interpretation.

# Works Cited

https://www.afrobarometer.org/wp-content/uploads/2022/02/tun_r7_sor_eng.pdf
